The Analog Method for Precipitation Prediction: Finding Better Analog Situations at a Sub-Daily Time Step

### 2016-11-24 Moving temporal window - HESS/template.tex
Corrections of the HESS paper.

@@

\begin{abstract}
The Analog Method (AM) aims at predicting local weather variables (predictands), such as precipitation, by means of a statistical relationship with predictors at a synoptic scale. The analogy is generally first assessed on the geopotential field gradients, in order to sample days with a similar atmospheric circulation. Other predictors, such as moisture variables, can also be added in a successive level of analogy.

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day, for both the target day and the candidate analogs. The constraint is the use of daily time series, due to the length of available archives they provide, and the unavailability of equivalent archives at a finer time step. However, it is unlikely that the best analogy happens at the very same hour, but it may occur at a different time of the day. In order to assess the potential of finding better analogs at a different hour, a moving time window (MTW) has been introduced on a reduced archive of hourly precipitation totals.

The MTW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criterion on the whole distribution of analog dates. The improvement was found to grow with the analog ranks due to an accumulation of more similar situations in the selection. Moreover, the improvement was even more important for days with heavy precipitation events, which are generally related to more dynamic atmospheric situations, where timing is more specific, and which are less present in the archive.

A seasonal effect has also been identified, with larger improvements in winter than in summer, supposedly due to the stronger effect of the diurnal cycle in summer, which favors predictors at the same hour for target and analogs.

The impact of the MTW on the prediction performance has been assessed by means of a sub-daily precipitation series transformed into moving 24~h-totals at a 6-hourly time step. This resulted in an improvement of the prediction skills, which were even larger after recalibrating the AM parameters.

However, the MTW requires sub-daily precipitation time series, which are generally available on a shorter period than daily precipitation archives. Thus, two simple temporal disaggregation methods were assessed in order to make longer archives usable with the MTW. The assessed approaches were not successful, which emphasized the need to use time series with an appropriate chronology. These should be available in a near future, either by means of growing observed archives, or by the establishment of precipitation reanalyses through regional modeling. Then, the use of a MTW in the AM can be considered for any application, and in any context, let it be operational forecasting or climate-related studies

@@

Analog methods (AMs) are based on the hypothesis that two relatively similar synoptic situations may produce similar local effects \citep{Lorenz1956, Lorenz1969}. They are part of statistical downscaling methods \cite{Maraun2010}. They consist in finding past situations that are similar, in terms of the atmospheric circulation or other synoptic predictors, to the target day of interest. Then, the local weather variables of interest (predictand) that were observed at the analog dates are used to construct a probabilistic prediction for the target day \citep{Duband1970, Zorita1999}. Multiple variations of the methods exist \citep[a non exhaustive listing can be found in][]{BenDaoud2016}. AMs are used for operational precipitation forecasting, either in the context of flood forecasting or hydropower production \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Desaint2008a, GarciaHernandez2009b, Bliefernicht2010, Marty2010, Marty2012, Horton2012, Obled2014, Hamill2015, BenDaoud2016}, as well as precipitation downscaling in a climate perspective \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015}. Other applications focus on temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectories of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

The spatial transferability of the method is analyzed in \citet{Chardon2014}. A great advantage of AMs is that they create realistic events, as they are based on observed situations, with consistent spatial distribution, as long as the analog dates chosen for a region are the same. The temporal transferability has also been the topic of recent studies, may it be for past or future climates \citep{Dayon2015, Caillouet2016}.

\citet{Ruosteenoja1988} and \citet{Vandendool1994} analyze the influence of the archive length on the quality of the analogy. They highlight a three-way relationship between the quality of the analogy, the archive length and the size of the spatial domain (of degrees of freedom): errors increase with a bigger domain, but decrease with a longer archive. They demonstrate that it is not possible to find good analogues on the whole northern hemisphere with a 100 years archive (and even for much higher orders of magnitude). For that reason, smaller spatial windows are always considered to search for analogues, and the archive length is maximized.

@@

The MTW allows finding analog situations at different hours of the day, which can also be seen as an inflation of the archive length. However, despite having $x$ time more candidate situations, the quantity of new information is not expected to be as important as a $x$ time longer archive, due to significant correlation between successive situations. Nevertheless, if the MTW can improve the prediction skill of the AM, it means that it does bring some new information from the archive. Thus, if the reduction of the archive length necessary to the MTW, due to the reduced availability of sub-daily precipitation time series, is expected to decrease the AM performance, the archive inflation brought by the MTW is expected to contribute to an increase in performance.

Other possibilities exist in order to increase the prediction skill of AMs. A classical approach is to add new predictors or new successive levels of analogy \citep[see e.g.][]{Horton2012a, BenDaoud2016, Caillouet2016}. AMs can also be combined with other methods \citep[see e.g.][]{Chardon2014}. Another possibility is to use a global optimization technique, such as genetic algorithms, in order to better optimize the method and to add new options \citep{Horton2016}. However, the MTW technique is not in competition with other approaches, but can be combined with these. Indeed, as most of them rely on the atmospheric circulation in the first level of analogy, the application of a MTW should lead to similar conclusions.

Section \ref{sec:data_methods} presents the context of the study as well as data and methods, including the proposed MTW technique. The impact of the reduction of the archive and the improvements brought by the MTW are assessed in Sect. \ref{sec:results}. The benefit of introducing a MTW was assessed first in regards to the analogy criterion improvement (Sect. \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (Sect. \ref{sec:influence_scores}). Finally, the results are discussed in Sect. \ref{sec:discussion}.

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland. Precipitation time series come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Fig. \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The data were available at an hourly time step on 25~years (1982 to 2007), and were also obtained at a daily time step on 47~years (1961 to 2008). Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The results will thereafter be presented arbitrarily for the Ulrichen station, but were equivalent for all stations.

Synoptic-scale variables, used as predictors, were extracted from the NCEP/NCAR reanalysis I \citep{Kalnay1996} dataset with a 6-hourly temporal resolution, 17 pressure levels, and a spatial resolution of 2.5\degree. This dataset is now relatively old, but still often used (see discussion in Sect. \ref{sec:old_reanalysis}).

@@

The predictand, here a 24~h precipitation total (classically from 6~h~UTC to 6~h~UTC the next day), results from meteorological situations that are continuous transitions between changing circulation patterns. Thus, the predictor data, extracted from reanalysis datasets, are considered at different hours of the day. Based on \citet{Bontron2005}, geopotential heights are thus compared at 1000~hPa (Z1000) at 12~h~UTC and 500~hPa (Z500) at 24~h~UTC. The time of the day at which the predictors are selected is found by \citet{Bontron2004} to have a significant influence.

@@

The target situations and their corresponding observed precipitation values do not change, because the prediction is always established for a fixed period of the target day (6~h~UTC to 6~h~UTC the next day), as before. The difference is that the candidates are 4~times more numerous (even though they are not fully independent). No constraint was added in order to restrict the selection of multiple analogs within the same candidate dates.

In order to assess the benefit of searching analog situations at a sub-daily time step, an appropriate precipitation series is required. On the basis of high resolution time series (Sect. \ref{sec:data}), 24~h-totals were processed, but at a 6-hourly time step, by means of a moving 24~h-total.  As sub-daily precipitation time series are available on a shorter period than daily ones, the loss of performance resulting from a reduction of a 47~years archive (1961 to 2008) to 25~years (1982 to 2007) needs to be assessed.

@@

It would be profitable to be able to apply the MTW to a longer archive (here, 1961 to 2008), rather than being limited to the reduced period corresponding to the sub-daily precipitation data. Thus, in order to reconstruct a longer archive of moving 24~h-totals, two simple disaggregation approaches of the daily precipitation time series were assessed.

@@

The second approach aimed at getting closer to the chronology of the actual precipitation by relying on some informative variable during the reconstruction procedure. The best proxy variable was identified by means of a correlation analyses (on non-zero values) with the 6-hourly precipitation time series on the period 1982-2007. Precipitable water and relative humidity (at 1000~hPa, 925~hPa or 850~hPa) were assessed at the four points surrounding the catchment, and by allowing a time lapse between both series in order to take into account the significant distance separating the weather stations and the reanalysis grid point. Data from the NCEP/NCAR reanalysis 1 (Sect. \ref{sec:data}) was used in this perspective, despite their rough resolution. Then, once the best proxy has been selected, its temporal profile was used in order to disaggregate the precipitation time series on the long archive period.

@@

To calibrate the method, the semi-automatic sequential procedure elaborated by \citet{Bontron2004} was used \cite[see also ][]{Horton2012a, Radanovics2013, BenDaoud2016}. The analogy levels (eg. the atmospheric circulation or moisture index) are calibrated sequentially. The parameters calibrated by this approach, for every level of analogy, are the spatial windows on which the predictors are compared and the number of analogs. The procedure consists of the following steps \citep{Bontron2004}:

@@

	\item For every level of analogy:
	\begin{enumerate}
		\item Identification of the most skilled unitary cell (1~point for moisture variables and 4 for the geopotential height when using the S1 criterion) over a large domain. Every point (or cell) of the full domain is assessed on all predictors of the level of analogy jointly (consisting generally of the same variable, but on different pressure levels and at different hours).
		\item From this most skilled point, the spatial window is expanded by successive iterations in the direction of greater performance gain. The spatial window grows up by repeating the previous steps, until no improvement is reached.
		\item The number of analog situations is then reconsidered and optimized for the current level of analogy.
	\end{enumerate}
	\item A new level of analogy can then be added, based on other variables (such as the moisture index) at chosen pressure levels and hours of the day. The number of analogs for the next level of analogy is initiated at a chosen value. Then, the procedure starts again from step (2) for the new level. The parameters calibrated on the previous analog levels are fixed and do not change. 
	\item Finally, the numbers of analogs are re-assessed for the different levels of analogy. This is done iteratively by varying the number of analogs of each level in a systematic way.
\end{enumerate}

@@

The performance loss resulting from a reduction of a 47~years archive (1961 to 2008, corresponding to the daily time series) to 25~years (1982 to 2007, corresponding to the hourly time series, see Sect. \ref{sec:mtw}) was assessed with the original method, without MTW. Both 2Z (Table \ref{table:method_2Z}) and 2Z-2MI (Table \ref{table:method_2Z-2MI}) methods were considered. The AM parameters were calibrated on the original archive (see resulting analog numbers in Tables \ref{table:analog_nb}) and were be used thereafter.

@@

\subsubsection{Changes in the atmospheric circulation analogy}

When searching for analogs in the first level of analogy, such as on the geopotential heights in the 2Z method (Sect. \ref{sec:analog_method}), there are 4 times more candidates (even though not fully independent) with the MTW than before (Sect. \ref{sec:mtw}). Figure \ref{fig:changes_S1_analogs} shows the changes in the distributions of the analogy criterion (S1) for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analog rank at the Ulrichen station on the whole calibration period (1961-2008 -- the full period could be used here, as precipitation were not considered at this stage), due to introduction of the MTW. The shape of the distributions of the conventional approach and the MTW were found to be similar, but the values of the analogy criterion were reduced (shifted to the left), and therefore better. For analogs with higher ranks (e.g. $20^{th}$ or $40^{th}$), the difference between the two distributions was larger than for the first ranks, which means that the improvement increased with the rank of the analogs.

@@

The impact of the MTW on the analogy criterion has been analyzed per precipitation classes. The results are summarized in Fig. \ref{fig:changes_S1_precip_threshold} by the median reduction of S1 for days with precipitation (organized into classes) between two thresholds. The number of cases per class being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement were more important on days with higher precipitation. It was the same for all other stations.

@@

Atmospheric dynamics varies greatly from one season to another, which reflects on the performance of the AM that is generally lower between June and August \citep{Bliefernicht2010}. The effect of the MTW on the S1 criterion per season is presented in Fig. \ref{fig:changes_S1_seasons}. Differences between the seasons were substantial. The improvements were greater for winter than summer.

\subsubsection{Changes in the moisture analogy}

When adding the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidate situations did not increase, as they were subsampled in the previously $N_{1}$ selected analogs, but their dates have changed. In contrast to the AM on the atmospheric circulation only, both a reduction or an increase of the RMSE analogy criterion values were possible here with the MTW compared to the static approach. Indeed, Fig. \ref{fig:changes_RMSE} shows an almost insignificant improvement of the RMSE values. Unlike for the first level of analogy, the relative improvement of the RMSE values were distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations were similar.

@@

The S1 criterion was previously found to be improved to a greater extent for higher precipitation values (Sect. \ref{sec:influence_precip}). The changes in terms of performance scores were then also assessed per precipitation classes. Fig. \ref{fig:changes_CRPS_precip_threshold} synthesizes these differences for the Ulrichen station, other stations having the same behavior. The performance score was improved for days with higher precipitation after the introduction of the MTW. On the contrary, non-rainy days and small precipitation values were not improved.

@@

After recalibrating, changes in the AM parameters could be observed for both 2Z and 2Z-2MI methods. Among these, the zonal extent of the spatial windows of the circulation analogy decrease a bit (not shown). More significantly, the optimal number of analogs increased after introducing the MTW, of a considerable order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method (Table \ref{table:analog_nb}). The number of analogs of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station.

@@

For the second approach, the synoptic proxy to use for the temporal disaggregation of the daily time series had to be identified first. The results are illustrated for the Zermatt station. Amongst the considered moisture variables (see Sect. \ref{sec:reconstruction}), the best proxy was the precipitable water at 45° N and 7.5° E, with a time offset of -6~h (Table \ref{table:proxy_correlations}). Table \ref{table:proxy_CRPSS} presents the CRPSS scores obtained by the reconstituted series using the proxy of precipitable water.

@@

On the first level of analogy, the median of the S1 values reduction (Fig. \ref{fig:changes_S1} bottom) started approximately at 5~\% for the first analog and reached more than 10~\% for the last one. This increase of improvements with the analog rank can be explained by the accumulation of better analogs in the distribution. The curve representing the minimum improvement is quickly slightly superior to zero, meaning that the criteria have been improved on most analogs for every day of the calibration period. All other stations had a similar improvement of the S1 criterion, both in terms of distribution shape and amplitude. The introduction of the MTW allowed finding better analog situations in the first level of analogy, resulting in a selection of days with more similar atmospheric circulation.

When adding a second level of analogy on moisture variables, the criteria values (RMSE) were not improved by the MTW as the number of candidates was not higher (still the $N_{1}$ days selected in the first level of analogy). However, this result of a globally null improvement of the RMSE values does not mean that the 2Z-2MI method cannot be improved by the MTW. It means that after the selection of the analog situations in terms of the synoptic circulation, the new candidate dates did not allow to find better analogs in terms of moisture. However, the selected dates have changed in the first level of analogy, and thus also in the final selection, which can potentially improve the prediction skill.

@@

Finding better analog situations does not necessarily imply better skills to predict the precipitation. This had to be assessed and the MTW was found to improve the skill of the precipitation prediction for both the 2Z and 2Z-2MI methods (Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}). Moreover, it required no additional predictor. With the introduction of the MTW, the performance loss related to the reduction of the archive length was compensated. It was here of the same order of magnitude as if the length of the archive doubled. Note that despite the number of candidate situations were four times higher, the gains seem to be lower than for a quadrupled archive length. The likely reason is that an actual longer archive would contain more atmospheric situations that might have been less observed on a shorter period. Moreover, high correlations between sub-daily circulation patterns are expected. The MTW thus did not create four times more independent data, but the increase in information was nevertheless substantial. In this regards, a reanalysis dataset with a higher temporal resolution might not improve the performance significantly more than a 6-hourly dataset. On the other hand, reducing the MTW to a 12-hourly time step might reduce redundancy in the archive, but might also reduce the chance of finding better analog situations, as circulation patterns can evolve substantially in 12~hours.

@@

Both the analogy criteria (Sect. \ref{sec:influence_precip}) and the performance scores (Sect. \ref{sec:improvement_CRPSS_precip_threshold}) were improved to a greater extent for heavier precipitation events. This is likely due to the fact that higher precipitation events are a consequence of atmospheric conditions with a greater dynamics, such as weather disturbances, which have a well marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure center and the fronts change significantly during a day. Moreover, these situations are less numerous than anticyclonic situations, which makes it less likely to find very good analogs at the same time of the day. We can therefore expect to improve more significantly these situations with a greater dynamics when introducing a MTW, as better matches to the target situation may be found. On the opposite, days with a low dynamics of the atmospheric circulation, such as anticyclonic situations, will not be radically improved by the introduction of the MTW.

>>>>

%TODO: check figs and table captions
%TODO: order figs
%TODO: english editing
%TODO: analogs
%TODO: perspective Obled: MTW without precip -> indicator

@@

\begin{abstract}
The Analog Method (AM) aims at predicting local weather variables (predictands), such as precipitation, by means of a statistical relationship with predictors at a synoptic scale. The analogy is generally first assessed on gradients of geopotential fields, in order to sample days with a similar atmospheric circulation. Other predictors, such as moisture variables, can also be added in a successive level of analogy.

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day, for both the target day and the candidate analogs. The main reason is the use of daily precipitation time series, due to the length of their available archives, and the unavailability of equivalent archives at a finer time step. However, it is unlikely for the best analogy to happen at the very same hour, as it may occur at a different time of the day. In order to assess the potential of finding better analogs at a different hour, a moving time window (MTW) has been introduced on a reduced archive of hourly precipitation totals.

The MTW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criterion on the whole distribution of analog dates. The improvement was found to grow with the analog rank due to an accumulation of more similar situations in the selection. A seasonal effect has also been identified, with larger improvements in winter than in summer, supposedly due to stronger diurnal cycles in summer, which favor predictors at the same hour for the target and analog days.

The impact of the MTW on the prediction skill has been assessed by means of a sub-daily precipitation series transformed into moving 24~h-totals at a 6-hourly time step. The prediction skill was found to be improved by the MTW, and even more after recalibrating the AM parameters. Moreover, the improvement was greater for days with heavy precipitation, which are generally related to more dynamic atmospheric situations, where timing is more specific, and which are less present in the archive.

However, the MTW requires sub-daily precipitation time series, which are generally available on a shorter period than daily precipitation archives. Thus, two simple temporal disaggregation methods were assessed in order to make longer archives usable with the MTW. The assessed approaches were not successful, which emphasized the need to use time series with an appropriate chronology. These should be available in a near future, either by means of growing archives of measurements, or by the establishment of regional precipitation reanalyzes. Then, the use of a MTW in AMs can be considered for any application, and in any context, let it be operational forecasting or climate-related studies.

@@

Analog methods (AMs) are based on the hypothesis that two relatively similar synoptic situations may produce similar local effects \citep{Lorenz1956, Lorenz1969}. They are part of statistical downscaling methods \citep{Maraun2010}, and consist in finding past situations that are similar, in terms of the atmospheric circulation or other synoptic predictors, to the target day of interest. Then, the local weather variables of interest (predictand) that were observed at the analog dates are used to construct a probabilistic prediction for the target day \citep{Duband1970, Zorita1999}. Multiple variations of the method exist \citep[a non-exhaustive listing can be found in][]{BenDaoud2016}. AMs are used for operational precipitation forecasting, either in the context of flood forecasting or hydropower production \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Desaint2008a, GarciaHernandez2009b, Bliefernicht2010, Marty2010, Marty2012, Horton2012, Obled2014, Hamill2015, BenDaoud2016}, as well as precipitation downscaling in a climate perspective \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015}. Other applications focus on temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectories of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

The spatial transferability of the method is analyzed in \citet{Chardon2014}. A great advantage of AMs is that they create realistic precipitation patterns, as they are based on observed situations, with consistent spatial distribution, as long as the analog dates chosen for a region are the same. The temporal transferability has also been the topic of recent studies, may it be for past or future climates \citep{Dayon2015, Caillouet2016}.

\citet{Ruosteenoja1988} and \citet{Vandendool1994} have analyzed the influence of the archive length on the quality of the analogy. They highlight a three-way relationship between the quality of the analogy, the archive length and the size of the spatial domain (or degrees of freedom): errors increase with a bigger domain, but decrease with a longer archive. They demonstrate that it is not possible to find good analogues on the whole northern hemisphere with a 100 years archive (and even for much higher orders of magnitude). For that reason, smaller spatial windows are always considered to search for analogues, and the archive length is maximized.

@@

The MTW allows finding analog situations at different hours of the day, which can also be seen as an inflation of the archive length. However, despite having $x$ time more candidate situations, the quantity of new information is not expected to be as important as a $x$ time longer archive, due to significant correlation between successive situations within the same day. Nevertheless, if the MTW can improve the prediction skill of the AM, it means that it does extract new information from the archive. Thus, if the reduction of the archive length necessary to the MTW, due to the reduced availability of sub-daily precipitation time series, is expected to decrease the AM performance, the archive inflation brought by the MTW is expected to contribute to an increase in performance.

Other possibilities exist in order to increase the prediction skill of AMs. A classical approach is to add new predictors or new successive levels of analogy \citep[see e.g.][]{Horton2012a, BenDaoud2016, Caillouet2016}. AMs can also be combined with other methods \citep[see e.g.][]{Chardon2014}. Another possibility is to use a global optimization technique, such as genetic algorithms, in order to better optimize the method and to add new parameters \citep{Horton2016}. However, the MTW technique is not in competition with other approaches, but can be combined with these. Indeed, as most of them rely on the atmospheric circulation in the first level of analogy, the application of a MTW should lead to similar conclusions.

Section \ref{sec:data_methods} presents the context of the study as well as data and methods, including the proposed MTW technique. The impact of the reduction of the archive and the improvements brought by the MTW are assessed in Sect. \ref{sec:results}. The benefit of introducing a MTW was assessed first in regards to the analogy criterion improvement (Sect. \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (Sect. \ref{sec:influence_scores}). Finally, the results are discussed in Sect. \ref{sec:discussion} and the conclusions are found in Sect. \ref{sec:conclusions}.

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland. Precipitation time series come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Fig. \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The data were available at an hourly time step on 25~years (1982 to 2007), and were also obtained at a daily time step (from 6~h~UTC to 6~h~UTC the next day) on 47~years (1961 to 2008). Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The results will thereafter be presented arbitrarily for the Ulrichen station, but were equivalent for all stations.

Synoptic-scale variables, used as predictors, were extracted from the NCEP/NCAR reanalysis I dataset \citep{Kalnay1996} with a 6-hourly temporal resolution, 17 pressure levels, and a spatial resolution of 2.5\degree. This dataset is now relatively old, but it is not expected to affect the conclusions of the present study (see discussion in Sect. \ref{sec:old_reanalysis}).

@@

The predictor variables, extracted from reanalysis datasets, are considered at different hours of the day. Based on \citet{Bontron2005}, geopotential heights are thus compared at 1000~hPa (Z1000) at 12~h~UTC and 500~hPa (Z500) at 24~h~UTC. The time of the day at which the predictors are selected is found by \citet{Bontron2004} to have a significant influence.

@@

The target situations and their corresponding observed precipitation values do not change, because the prediction is still established on a daily basis, for a fixed period of the target day (6~h~UTC to 6~h~UTC the next day), as before. The difference is that the candidates are 4~times more numerous (even though they are not fully independent). No constraint was added in order to restrict the selection of multiple analogs within the same candidate dates.

In order to assess the benefit of searching analog situations at a sub-daily time step, an appropriate precipitation series is required. On the basis of high temporal resolution time series (Sect. \ref{sec:data}), 24~h-totals were processed, but at a 6-hourly time step, by means of a moving 24~h-total.  As sub-daily precipitation time series are available on a shorter period than daily ones, the loss of performance resulting from a reduction of a 47~years archive (1961 to 2008) to 25~years (1982 to 2007) needs to be assessed.

@@

It would be profitable to be able to apply the MTW to a longer archive (here, 1961 to 2008), rather than being limited to the reduced period corresponding to the sub-daily precipitation data (cf. previous section). Thus, in order to reconstruct a longer archive of moving 24~h-totals, two simple disaggregation approaches of the daily precipitation time series were assessed.

@@

The second approach aimed at getting closer to the chronology of the actual precipitation by relying on some informative variables during the reconstruction procedure. Data from the NCEP/NCAR reanalysis 1 (Sect. \ref{sec:data}) were used in this perspective, despite their rough resolution. Precipitable water and relative humidity (at 1000~hPa, 925~hPa or 850~hPa) were assessed at the four points surrounding the catchment, and by allowing a time lapse between both series in order to take into account the significant distance separating the weather stations and the reanalysis grid point. The best proxy variable was identified by means of a correlation analyses (on non-zero values) with the 6-hourly precipitation time series on the period 1982-2007. Then, once the best proxy has been selected, its temporal profile was used in order to disaggregate the daily precipitation time series.

@@

The semi-automatic sequential procedure elaborated by \citet{Bontron2004} was used to calibrate the AM \cite[see also ][]{Radanovics2013, BenDaoud2016}. The analogy levels (e.g. the atmospheric circulation or moisture index) are calibrated sequentially. The parameters calibrated by this approach, for every level of analogy, are the spatial windows on which the predictors are compared and the number of analogs. The procedure consists of the following steps \citep{Bontron2004}:

@@

	\item For every level of analogy:
	\begin{enumerate}
		\item Identification of the most skilled unitary cell (1~point for moisture variables and 4 for the geopotential height when using the S1 criterion) over a large domain. Every point (or cell) of the full domain is assessed on all predictors of the level of analogy, jointly (consisting generally of the same variable, but on different pressure levels and at different hours).
		\item From this most skilled point, the spatial window is expanded by successive iterations in the direction of greater performance gain. The spatial window grows up by repeating the previous steps, until no improvement is reached.
		\item The number of analog situations is then reconsidered and optimized for the current level of analogy.
	\end{enumerate}
	\item A new level of analogy can then be added, based on other variables (such as the moisture index) at chosen pressure levels and hours of the day. The number of analogs for the next level of analogy is initiated at a chosen value. Then, the procedure starts again from step (2) for the new level. The parameters calibrated on the previous analog levels are fixed and do not change. 
	\item Finally, the numbers of analogs for the different levels of analogy are re-assessed. This is done iteratively by varying the number of analogs of each level in a systematic way.
\end{enumerate}

@@

The performance loss resulting from the reduction of a 47~years archive (1961 to 2008, corresponding to the daily time series) to 25~years (1982 to 2007, corresponding to the hourly time series, see Sect. \ref{sec:mtw}) was assessed with the original method, without MTW. Both 2Z (Table \ref{table:method_2Z}) and 2Z-2MI (Table \ref{table:method_2Z-2MI}) methods were considered. The AM parameters were calibrated on the original archive (see resulting analog numbers in Tables \ref{table:analog_nb}) and were used thereafter.

@@

\subsubsection{Changes in the atmospheric circulation analogy}
\label{sec:perf_2Z}

When searching for analogs in the first level of analogy, such as on the geopotential heights in the 2Z method (Sect. \ref{sec:analog_method}), there are 4 times more candidates (even though not fully independent) with the MTW than before (Sect. \ref{sec:mtw}). Figure \ref{fig:changes_S1_analogs} shows the changes in the distributions of the analogy criterion (S1) for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analog rank at the Ulrichen station on the whole calibration period (1961-2008 -- the full period could be used here, as precipitation were not considered at this stage), due to introduction of the MTW. The shape of the distributions of the conventional approach and the MTW were found to be similar, but the values of the analogy criterion were reduced (shifted to the left), and were therefore better. For analogs with higher ranks (e.g. $20^{th}$ or $40^{th}$), the difference between the two distributions was larger than for the first ranks, which means that the improvement increased with the rank of the analogs.

@@

The impact of the MTW on the analogy criterion has been analyzed per precipitation classes (for the target day). The results are summarized in Fig. \ref{fig:changes_S1_precip_threshold} by the median reduction of S1 for days with precipitation (organized into classes) between two thresholds. The number of cases per class being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement were more important on days with higher precipitation. Once again, the results for the other stations were similar.

@@

Atmospheric dynamics varies greatly from one season to another, which reflects on the performance of the AM that is generally lower between June and August \citep{Bliefernicht2010}. The effect of the MTW on the S1 criterion per season is presented in Fig. \ref{fig:changes_S1_seasons}. Differences between the seasons were substantial, with greater improvements for winter than summer.

\subsubsection{Changes in the moisture analogy}
\label{sec:perf_2Z-2MI}

When adding the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidate situations did not increase, as they were conditioned by the $N_{1}$ previously selected analogs, but their dates have changed. In contrast to the AM on the atmospheric circulation only, both a reduction or an increase of the RMSE analogy criterion values were possible here with the MTW compared to the static approach. Indeed, Fig. \ref{fig:changes_RMSE} shows an almost insignificant improvement of the RMSE values. Unlike for the first level of analogy, the relative changes of the RMSE values were distributed relatively symmetrically around zero, with improvements and losses of the same amplitude.

@@

The S1 criterion was previously found to be improved to a greater extent for higher precipitation values (Sect. \ref{sec:influence_precip}). The changes in terms of performance scores were then also assessed per precipitation classes. Fig. \ref{fig:changes_CRPS_precip_threshold} synthesizes these differences for the Ulrichen station, the other stations having the same behavior. The performance score was improved for days with higher precipitation after the introduction of the MTW. On the contrary, non-rainy days and small precipitation values were not improved.

@@

After recalibrating, changes in the AM parameters could be observed for both 2Z and 2Z-2MI methods. Among these, the zonal extent of the spatial windows of the circulation analogy decrease a bit (not shown). More significantly, the optimal number of analogs increased after introducing the MTW, of a considerable order of magnitude: 25~\% to 83~\% for the 2Z method, and 20~\% to 67~\% for the final selection of the 2Z-2MI method (Table \ref{table:analog_nb}). The number of analogs of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station.

@@

For the second approach, the synoptic proxy to use for the temporal disaggregation of the daily time series had to be identified first. The results are illustrated this time for the Zermatt station. Amongst the considered moisture variables (see Sect. \ref{sec:reconstruction}), the best proxy was the precipitable water at 45° N and 7.5° E, with a time offset of -6~h (Table \ref{table:proxy_correlations}). Table \ref{table:proxy_CRPSS} presents the CRPSS scores obtained by the disaggregated series using the proxy of precipitable water.

@@

On the first level of analogy, the improvement of the S1 criterion (Fig. \ref{fig:changes_S1} bottom) started approximately at 5~\% for the first analog and reached more than 10~\% for the last one (Sect. \ref{sec:perf_2Z}). This increase of improvements with the analog rank can be explained by the accumulation of better analogs in the selection. The curve representing the minimum improvement is quickly slightly superior to zero, meaning that the criteria have been improved on most analog ranks for every day of the calibration period. All other stations had a similar improvement of the S1 criterion, both in terms of the distribution shape and amplitude. The introduction of the MTW allowed finding better analog situations in the first level of analogy, resulting in a selection of days with more similar atmospheric circulation.

When adding a second level of analogy on moisture variables, the criteria values (RMSE) were not improved by the MTW (Sect. \ref{sec:perf_2Z-2MI}), as the number of candidates was not higher (still the $N_{1}$ days selected in the first level of analogy). However, this result of a globally null improvement of the RMSE values does not mean that the 2Z-2MI method cannot be improved by the MTW. It means that after the selection of the analog situations in terms of the synoptic circulation, the new candidate dates did not allow to find better analogs in terms of moisture. However, the selected dates have changed in the first level of analogy, and thus also in the final selection, which did improve the prediction skill (Sect. \ref{sec:influence_scores}).

@@

The MTW was found to improve the skill of the precipitation prediction for both the 2Z and 2Z-2MI methods (Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}, Sect. \ref{sec:influence_scores}). Moreover, it required no additional predictor. With the introduction of the MTW, the performance loss related to the reduction of the archive length was compensated. It was here of the same order of magnitude as if the length of the archive doubled. Note that despite the number of candidate situations being four times higher, the gains seem to be lower than for a quadrupled archive length. The likely reason is that an actual longer archive would contain more atmospheric situations that might have been less observed on a shorter period. Moreover, high correlations between sub-daily circulation patterns are expected. The MTW thus did not create four times more independent data, but the increase in information was nevertheless substantial. In this regards, a reanalysis dataset with a higher temporal resolution might not improve the performance significantly more than a 6-hourly dataset. On the other hand, reducing the MTW to a 12-hourly time step might reduce redundancy in the archive, but might also reduce the chance of finding better analog situations, as circulation patterns can evolve substantially in 12~hours.

@@

Both the analogy criteria (Sect. \ref{sec:influence_precip}) and the performance scores (Sect. \ref{sec:improvement_CRPSS_precip_threshold}) were improved to a greater extent for days with heavier precipitation. This is likely due to the fact that higher precipitation events are a consequence of atmospheric conditions with greater dynamics, such as weather disturbances, which have a well-marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure center and the fronts change significantly during a day. Moreover, these situations are less numerous than anticyclonic situations, which makes it less likely to find very good analogs at the same time of the day. We can therefore expect to improve more significantly these situations with greater dynamics when introducing a MTW, as better matches to the target situation may be found. On the opposite, days with low dynamics of the atmospheric circulation, such as anticyclonic situations, will not be radically improved by the introduction of the MTW.
